Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-11-15 ms.tex
Phil's last changes to the appendix

The radiation transfer equation for resonant scattering is solved in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with  mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct boundary condition of zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of the correction is order unity when the spectral peaks approach the Doppler core and decreases with line center optical depth, which may explain discrepancies in previous studies. For the impulsive problem, the time, spatial, and frequency dependence of the equation are expressed using an eigenfunction expansion and are obtained numerically to characterize the escape time distribution and emergent spectra of photons. It is shown that the lowest-order eigenfrequency of the semi-analytic calculation agrees well with the decay constant of the Monte Carlo escape time distribution at sufficiently large line-center optical depths.

>>>>

The radiation transfer equation for resonant scattering is solved in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with  mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct boundary condition of zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of the correction is order unity when the spectral peaks approach the Doppler core and decreases slowly with line center optical depth, which may explain discrepancies in previous studies. For the impulsive problem, the time, spatial, and frequency dependence of the solution are expressed using an eigenfunction expansion in order to characterize the escape time distribution and emergent spectra of photons. It is shown that the lowest-order eigenfrequency agrees well with the decay rate found in the Monte Carlo escape time distribution at sufficiently large line-center optical depths.
